---
title: "Voter EDA"
author: "Keith G. Williams 800690755"
date: "Tuesday, April 19, 2016"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

This document is meant to be a mock-up of ideas for DSBA 5122 Visual Analytics Project.

Load the Voter Data from Mecklenburg County Board of Elections.  
```{r, cache=TRUE}
library(readr)
setwd('~/DSBA 5122/Project')

voter <- read_csv('meck_voter.csv')
```

Clean  
```{r}
library(stringr)

# Party Codes
voter$party_code <- str_replace_all(voter$party_code, '\r', '')
voter$party_code[voter$party_code == ""] <- NA

# District Descriptions
voter$ward_desc <- str_replace_all(voter$ward_desc,
                                   'CITY COUNCIL DISTRICT ',
                                   '')
voter$cong_dist_desc <- str_replace_all(voter$cong_dist_desc,
                                   'CONGRESSIONAL DISTRICT ',
                                   '')
voter$super_court_desc <- str_replace_all(voter$super_court_desc,
                                   'JUDICIAL DISTRICT ',
                                   '')
voter$nc_senate_desc <- str_replace_all(voter$nc_senate_desc,
                                   'NC SENATE DISTRICT ',
                                   '')
voter$nc_house_desc <- str_replace_all(voter$nc_house_desc,
                                   'NC HOUSE DISTRICT ',
                                   '')
voter$county_commiss_desc <- str_replace_all(voter$county_commiss_desc,
                                   'BOARD OF COMMISSIONERS DISTRICT ',
                                   '')
voter$school_dist_desc <- str_replace_all(voter$school_dist_desc,
                                   'SCHOOL BOARD DIST ',
                                   '')
voter$judic_dist_desc <- str_replace_all(voter$judic_dist_desc,
                                   'JUDICIAL DISTRICT ',
                                   '')
```

Visualize vote counts by election date  
```{r, echo=FALSE}
library(dplyr)
library(ggplot2)

ct_by_date <- voter %>%
    group_by(edate, party_code) %>%
    summarise(n = n())

p1 <-ggplot(ct_by_date, aes(as.factor(edate), n)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle=90, vjust=0.5)) +
    labs(x = "Election Date",
         y = "Number of Votes")
p1
```

Examples of how this data can be "drilled-down". Imagine a Check box on the left that allows the user to select which attributes the counts should be aggregated by. When clicking on "Party", the previous graph turns into this graph.
```{r, echo=FALSE}
p2 <- ggplot(ct_by_date, 
       aes(as.factor(edate), n, fill=party_code)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values=c('#377eb8', '#4daf4a', '#e41a1c', '#969696')) +
    theme(axis.text.x = element_text(angle=90, vjust=0.5)) +
    labs(x = "Election Date",
         y = "Number of Votes")

p2
```

Since "LIB" appears so little, maybe the user can filter out these records:  
```{r, echo=FALSE}
p3 <- ggplot(filter(ct_by_date, party_code != "LIB"), 
    aes(as.factor(edate), n, fill=party_code)) +
    geom_bar(stat="identity", position = "dodge") +
    scale_fill_manual(values=c('#377eb8', '#e41a1c', '#969696')) +
    theme(axis.text.x = element_text(angle=90, vjust=0.5)) +
    labs(x = "Election Date",
         y = "Number of Votes")

p3
```

Example of making this interactive. Hover and Zoom features are nice here. Clicking on the legend is not that useful in this case.  
```{r, echo=FALSE}
library(plotly)
ggplotly(p2)
```

Other aggregations could be race, gender, age, congressional district (there are only 3 in Mecklenburg County), superior court district (again, only 3 in Mecklenburg County).

Next, a map like the one we can make in Tableau would be nice.

```{r, echo=FALSE}
# install from github
#library(devtools)
#install_github('arilamstein/choroplethrZip@v1.5.0')
library(choroplethrZip)
meck_county <- 37119 #FIPS code for Mecklenburg County

# create data frame for plotting
county_votes <- voter %>%
    group_by(zip_code) %>%
    summarise(value = n()) %>%
    rename(region = zip_code) %>%
    mutate(region = as.character(region))

zip_votes <- zip_choropleth(county_votes,
                            county_zoom = meck_county,
                            title = "Total Votes by Zip Code 2008 - 2015",
                            legend = "Number of Votes",
                            num_colors=1,
                            reference_map = FALSE)

zip_votes
```

This map could come with an option to filter on election dates or by party or by demographics. Here's an example with a reference map turned "ON":  
```{r, echo=FALSE}
zip_choropleth(county_votes, county_zoom = meck_county,
               title = "Total Votes by Zip Code 2008 - 2015",
               legend = "Number of Votes",
               num_colors=1,
               reference_map = TRUE)
```